---
title: "Curling Project - S&DS 4250"
output: pdf_document
author: "Jake Todd and Sonam Wangchuk"
---

```{r}
Competition <- read.csv("clean_data/Competition.csv")
Competitors <- read.csv("clean_data/Competitors.csv")
Ends <- read.csv("clean_data/Ends.csv")
Games <- read.csv("clean_data/Games.csv")
Stones <- read.csv("clean_data/Stones.csv")
Teams <- read.csv("clean_data/Teams.csv")

library(dplyr)
library(randomForest)
```

```{r}
Stones <- Stones %>%
  mutate(ShotID = recode(ShotID,
                    `7` = 1,
                    `8` = 2,
                    `9` = 3,
                    `16` = 4,
                    `17` = 5,
                    `18` = 6,
                    `19` = 7,
                    `20` = 8,
                    `21` = 9,
                    `22` = 10))
```

Cleaning out the 0 and the 4950s to have ShotYet and OutOfPlay indicators

```{r}
ShotYet <- function(Xvalue, Yvalue) {
  return(!(Xvalue == 0 & Yvalue == 0)) 
}

for (i in 1:12) {
  Stones[[paste0("Stone_", i, "_Shot")]] <-
    ShotYet(Stones[[paste0("stone_", i, "_x")]],
                      Stones[[paste0("stone_", i, "_y")]])
}

OutOfPlay <- function(Xvalue, Yvalue) {
  return(Xvalue == 4095 | Yvalue == 4095) 
}

for (i in 1:12) {
  Stones[[paste0("Stone_", i, "_OutOfPlay")]] <-
    OutOfPlay(Stones[[paste0("stone_", i, "_x")]],
                      Stones[[paste0("stone_", i, "_y")]])
}
```

Remove the wrong values and replace them with NA

```{r}

Stones <- Stones %>%
  mutate(across(all_of(paste0("stone_", 1:12, "_x")), 
                ~na_if(., 0)))
Stones <- Stones %>%
  mutate(across(all_of(paste0("stone_", 1:12, "_x")), 
                ~na_if(., 4095)))

Stones <- Stones %>%
  mutate(across(all_of(paste0("stone_", 1:12, "_y")), 
                ~na_if(., 0)))
Stones <- Stones %>%
  mutate(across(all_of(paste0("stone_", 1:12, "_y")), 
                ~na_if(., 4095)))
```

Calculate a bunch of other features about the stones

```{r}
EuclideanDistance <- function(Xvalue, Yvalue) {
  final <- sqrt((Xvalue - 750)^2 + (Yvalue - 800)^2)
  return(final) 
}

for (i in 1:12) {
  Stones[[paste0("Stone_", i, "_Distance")]] <-
    EuclideanDistance(Stones[[paste0("stone_", i, "_x")]],
                      Stones[[paste0("stone_", i, "_y")]])
}

Behind <- function(Yvalue) {
  return(Yvalue < 800) 
}

for (i in 1:12) {
  Stones[[paste0("Stone_", i, "_Behind")]] <-
    Behind(Stones[[paste0("stone_", i, "_y")]])
}

Left <- function(Xvalue) {
  return(Xvalue < 750) 
}

for (i in 1:12) {
  Stones[[paste0("Stone_", i, "_Left")]] <-
    Left(Stones[[paste0("stone_", i, "_x")]])
}
```

Transform the data to look at individual ends from team 1's perspective

```{r}
Ends_Team1 <- Ends %>%
  left_join(Games %>% 
              select(CompetitionID, SessionID, GameID, TeamID1, TeamID2),
            by = c("CompetitionID", "SessionID", "GameID")) %>%
  mutate(Is_Team1 = (TeamID == TeamID1)) %>%
  group_by(CompetitionID, SessionID, GameID, EndID) %>%
  summarize(
    TeamID1 = first(TeamID1),
    TeamID2 = first(TeamID2),
    Result_Team1 = sum(ifelse(Is_Team1, Result, -Result)),
    .groups = "drop"
  )

Stones <- Stones %>%
  left_join(Ends_Team1,
            by = c("CompetitionID", "SessionID", "GameID", "EndID"))
```

Adding a few more indicators. House assumes that the radius is about 6 feet = 600 units. This could be wrong?

One of the most important features to add is the hammer indicator. The team with the hammer throws the last stone in the end, which is a significant advantage.

```{r}
Stones <- Stones %>%
  rowwise() %>%
  mutate(
    Team1_In_House = sum(c_across(paste0("Stone_", 1:6, "_Distance")) < 600, na.rm = TRUE),
    Team2_In_House = sum(c_across(paste0("Stone_", 7:12, "_Distance")) < 600, na.rm = TRUE),

    Team1_Closest = {
      vals <- c_across(paste0("Stone_", 1:6, "_Distance"))
      if(all(is.na(vals))) NA_real_ else min(vals, na.rm = TRUE)
    },
    Team2_Closest = {
      vals <- c_across(paste0("Stone_", 7:12, "_Distance"))
      if(all(is.na(vals))) NA_real_ else min(vals, na.rm = TRUE)
    },
    Team1_Has_Closest = Team1_Closest < Team2_Closest,

    Team1_Guards = sum(c_across(paste0("stone_", 1:6, "_y")) > 1200, na.rm = TRUE),
    Team2_Guards = sum(c_across(paste0("stone_", 7:12, "_y")) > 1200, na.rm = TRUE),
    
    Current_Score_Team1 = ifelse(Team1_Has_Closest, 
                                  sum(c_across(paste0("Stone_", 1:6, "_Distance")) < Team2_Closest, na.rm = TRUE),
                                  -1 * sum(c_across(paste0("Stone_", 7:12, "_Distance")) < Team1_Closest, na.rm = TRUE))
  ) %>%
  ungroup() %>%
  mutate(
    Stones_On_Ice = rowSums(!is.na(select(., Stone_1_Distance:Stone_12_Distance)))
  ) 

Stones <- Stones %>%
  group_by(CompetitionID, SessionID, GameID, EndID) %>%
  mutate(
    Team1_Hammer = case_when(
      any(Stone_2_Shot == TRUE & Stone_8_Shot == FALSE, na.rm = TRUE) ~ FALSE,
      any(Stone_2_Shot == FALSE & Stone_8_Shot == TRUE, na.rm = TRUE) ~ TRUE,
      TRUE ~ NA
    )
  ) %>%
  ungroup() %>%
  group_by(CompetitionID, SessionID, GameID, EndID, TeamID) %>%
  mutate(
    AveragePoints = mean(Points, na.rm = TRUE)
  ) %>%
  ungroup()
```


```{r}
Stones <- Stones %>%
  group_by(CompetitionID, SessionID, GameID, EndID) %>%
  mutate(
    AveragePointsDifference = mean(Points[TeamID == unique(TeamID)[1]], na.rm = TRUE) - 
                             mean(Points[TeamID == unique(TeamID)[2]], na.rm = TRUE),
    
    # Create AverageInHouseDistance
    AverageInHouseDistance = {
      # Get all Stone_X_Distance columns
      stone_distances <- c(Stone_1_Distance, Stone_2_Distance, Stone_3_Distance, 
                          Stone_4_Distance, Stone_5_Distance, Stone_6_Distance,
                          Stone_7_Distance, Stone_8_Distance, Stone_9_Distance,
                          Stone_10_Distance, Stone_11_Distance, Stone_12_Distance)

      in_house_distances <- stone_distances[stone_distances < 600 & !is.na(stone_distances)]

      if (length(in_house_distances) > 0) {
        mean(in_house_distances, na.rm = TRUE)
      } else {
        NA_real_
      }
    }
  ) %>%
  ungroup()

```

########################### CLEANING COMPLETE ###########################

Just the relevant engineered data now:

```{r}
model_data <- Stones %>%
  filter(!is.na(Result_Team1)) %>% 
  select(
    Team1_In_House, Team2_In_House,
    Team1_Closest, Team2_Closest,
    Team1_Guards, Team2_Guards,
    Current_Score_Team1,
    ShotID, Team1_Hammer,
    Result_Team1,
    AverageInHouseDistance
  ) %>%
  na.omit() %>%
  mutate(
    Team1_Won_End = as.factor(ifelse(Result_Team1 > 0, "TRUE", "FALSE"))
  )
```

Let's try the most basic logistic regression model on this (I have no idea how to handle ties at the moment, so a tie also returns false):

```{r}
model <- glm(
  Team1_Won_End ~ Team1_In_House + Team2_In_House + 
    Team1_Closest + Team2_Closest + Team1_Guards + Team2_Guards + 
    Current_Score_Team1 + ShotID + Team1_Hammer * AverageInHouseDistance,
  data = model_data,
  family = binomial
)

summary(model)
```


```{r}
model_data$pred_prob <- predict(model, model_data, type = "response")
model_data$pred_class <- ifelse(model_data$pred_prob > 0.5, TRUE, FALSE)

accuracy <- mean(model_data$pred_class == model_data$Team1_Won_End, na.rm = TRUE)
accuracy
```

The following is a simple random forest model to predict whether Team 1 wins the end based on the engineered features above!

```{r}

set.seed(123)
train_idx <- sample(1:nrow(model_data), 0.8 * nrow(model_data))
train <- model_data[train_idx, ]
test <- model_data[-train_idx, ]

rf <- randomForest(
  Team1_Won_End ~ Team1_In_House + Team2_In_House + 
    Team1_Closest + Team2_Closest + Team1_Guards + Team2_Guards + 
    Current_Score_Team1 + ShotID + Team1_Hammer + AverageInHouseDistance,
  data = train,
  ntree = 200,
  importance = TRUE
)
test <- filter(test, ShotID > 2)
test$pred_prob <- predict(rf, test, type = "prob")[, "TRUE"]
sum(test$Team1_Won_End == ifelse(test$pred_prob > 0.5, "TRUE", "FALSE")) / nrow(test)
```

This one is more of a points prediction model. Accuracy is determined by the sign of the predicted score

```{r}
rf_score <- randomForest(
  Result_Team1 ~ Team1_In_House + Team2_In_House + 
    Team1_Closest + Team2_Closest + Team1_Guards + Team2_Guards + 
    Current_Score_Team1 + ShotID + Team1_Hammer + AverageInHouseDistance,
  data = train,
  ntree = 200,
  importance = TRUE
)
test <- filter(test, ShotID < 6)
test$pred_score <- predict(rf_score, test)
test$pred_score <- round(test$pred_score)
hist(test$pred_score - test$Result_Team1, breaks = 30, main = paste0("Prediction Errors. Accuracy: ", sum(test$pred_score == test$Result_Team1) / nrow(test)), xlab = "Predicted - Actual Score")
```

Scoring potential guards => Dead end

```{r}
Exp = Stones %>% filter(Task == 2, Points == 4)
```


```{r}
# Define mappings
map_hammer <- c(
  "1" = "stone_8",
  "2" = "stone_2",
  "3" = "stone_9",
  "4" = "stone_3",
  "5" = "stone_10",
  "6" = "stone_4",
  "7" = "stone_11",
  "8" = "stone_5",
  "9" = "stone_12",
  "10" = "stone_6"
)

map_no_hammer <- c(
  "1" = "stone_2",
  "2" = "stone_8",
  "3" = "stone_3",
  "4" = "stone_9",
  "5" = "stone_4",
  "6" = "stone_10",
  "7" = "stone_5",
  "8" = "stone_11",
  "9" = "stone_6",
  "10" = "stone_12"
)

# Select appropriate mapping for each row
stone_map <- ifelse(
  Exp$Team1_Hammer,
  map_hammer[as.character(Exp$ShotID)],
  map_no_hammer[as.character(Exp$ShotID)]
)

# Build column names for x and y
x_cols <- paste0(stone_map, "_x")
y_cols <- paste0(stone_map, "_y")

# Use mapply to pull the right columns row-by-row
Exp$Current_x <- mapply(function(row, col) Exp[row, col], seq_len(nrow(Exp)), x_cols)
Exp$Current_y <- mapply(function(row, col) Exp[row, col], seq_len(nrow(Exp)), y_cols)

```


```{r}
plot(Exp$Current_x, Exp$Current_y)
```


```{r}

```

